### 2024-06-06 products/manuscript/Manuscript.qmd
First Manuscript edit

---
title: "Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset "
subtitle: ""
author: Antonio Flores
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract
This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.

{{< pagebreak >}}

# Introduction

## General Background Information
Providing businesses with a model that will allow prioritization of consumers could save time and allow for a more accurate use of resources.

## Description of data and data source
The data was donated on 2/13/2012. It was collected from phone call marketing campaigns performed by a Portuguese banking institution.I have accessed this data from the UC Irvine Machine Learning Repository.

There are 45,212 records, and includes: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history. Additionally, the classification variable is whether or not the person subscribed to a term deposit. There are 16 features in total.

## Questions/Hypotheses to be addressed

The research question I plan to address with my analysis is which features or combination of features are the best predictors of consumers making a deposit. The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns. I don't currently have specific predictors in mind to begin with, as I plan to test all of the available features while also looking at some combinations or concatenation of features.

_To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:_

_Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]._

{{< pagebreak >}}

# Methods

I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already. Then I will test for correllation between different features and proceed from there.

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
```

## Data aquisition
Data retrieved from UCI ML Repository

## Data import and cleaning

### Reading in the Data

raw = read.csv("C:/Users/Client/Documents/Flores-P2-Project/bank-full.csv",header = TRUE,
                  sep=";")
head(raw)

```{r}
#| label: tbl-summarytable1
#| tbl-cap: "Data Snapshot"
#| echo: FALSE
raw=read.csv("../../bank-full.csv", header = TRUE, sep=";")
head(raw)
```

### Dimensions:
dim(raw)

```{r}
#| label: tbl-summarytable2
#| tbl-cap: "Data Dimensions"
#| echo: FALSE
dim(raw)
```

### Describing raw data
str(raw)

```{r}
#| label: tbl-summarytable3
#| tbl-cap: "Data Description"
#| echo: FALSE
str(raw)
```

## Statistical analysis
_Explain anything related to your statistical analyses._

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2024-06-14 code/eda-code/eda - my version.qmd
For exercise 2# git status git status

---
title: "An example exploratory analysis script"
date: "2024-02-07"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

I'm only showing it the way where the code is included in the file. 
As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.

But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)

Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.

Also note that while here I split cleaning and exploring, this is iterative. You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it. In general, as you explore, you'll find things that need cleaning. As you clean, you can explore more. Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file. Or split things in any other logical way.

As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data. Depending on the total number of variables in your dataset, explore all or some of the others. Figures produced here might be histograms or density plots, correlation plots, etc. Tables might summarize your data.

Start by exploring one variable at a time. Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in. If your dataset is small, you can do that for all variables.

Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(scales)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata1.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable1.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

Height first.

```{r}
p1 <- mydata %>% ggplot(aes(x=age)) + geom_histogram() 
plot(p1)
figure_file = here("results", "figures","age-distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Now weights.

```{r}
p2 <- mydata %>% ggplot(aes(x=balance)) + geom_histogram() + scale_x_log10(labels =label_dollar())
plot(p2)
figure_file = here("results", "figures","balance-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now height as function of weight.

```{r}
p3 <- mydata %>% ggplot(aes(x=age, y=balance)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results", "figures", "age-balance-scatter.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

```{r}
p4 <- mydata %>% ggplot(aes(x=age, y=balance, color = marital)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results", "figures", "age-balance-stratified.png")
ggsave(filename = figure_file, plot=p4) 
```

```{r}
p5 = mydata |>
  ggplot(aes(balance, job))+
  geom_boxplot()+
  scale_x_log10(labels =label_dollar())
p5
figure_file = here("results", "figures", "job-balance-scatter.png")
ggsave(filename = figure_file, plot=p5)

```

```{r}

mydata |>
  ggplot(aes(job))+
  geom_bar()

mydata |>
  ggplot(aes(education))+
  geom_bar()

mydata |>
  ggplot(aes(marital))+
  geom_bar()

```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

### 2024-06-14 code/processing-code/processingfile-v1 - my version.qmd
For exercise 2# git status git status

---
title: "An example cleaning script"
author: "Antonio Flores"
date: "2024-13-06"
output: html_document
---

# Processing script

This contains the same code and comments/information as `processingcode.R`.

This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.

See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) #for loading Excel files
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
```

# Data loading

Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so:
package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives",
but it adds typing. You can do it either way.

```{r}
# path to data
# note the use of the here() package and not absolute paths

data_location <- here::here("data","raw-data","bank-full.csv")
rawdata <- read.csv(data_location, header = TRUE, sep=";")
```

# Check data

First we can look at the codebook

```{r}
codebook_location <- 
here::here("data","raw-data","codebook.xlsx")

codebook <- readxl::read_excel(codebook_location, sheet ="Codebook")
print(codebook)
```

Several ways of looking at the data

```{r}
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning
```{r}

d1 = rawdata |> 
  mutate(job = as.factor(job)) |>
  mutate(marital = as.factor(marital)) |>
  mutate(education = as.factor(education)) |>
  mutate(default = as.factor(default)) |>
  mutate(housing = as.factor(housing)) |>
  mutate(loan = as.factor(loan)) |>
  mutate(education = as.factor(education)) |>
  mutate(contact = as.factor(contact)) |>
  mutate(month = as.factor(month)) |>
  mutate(poutcome = as.factor(poutcome)) |>
  mutate(y = as.factor(y))

summary(d1)

```

All done, data is clean now.

Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.

```{r}
processeddata1 <- d1
```

# Save data

Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.

See here for some suggestions on how to store your processed data:
http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata

```{r}
save_data_location <- here::here("data","processed-data","processeddata1.rds")
saveRDS(processeddata1, file = save_data_location)
```

# Notes

Removing anyone who had "faulty" or missing data is one approach. It's often not the best. based on your question and your analysis approach, you might want to do cleaning differently (e.g. keep individuals with some missing information).

### 2024-06-14 code/eda-code/eda - my version.qmd
no more template info

@@
 ---
-title: "An example exploratory analysis script"
+title: "Exploratory Data Analysis script"
 date: "2024-02-07"
 output: html_document
+author: Antonio Flores
 ---
@@
 
-I'm only showing it the way where the code is included in the file. 
-As described in the `processing_code` materials, I currently prefer the approach of having R code in a separate file and pulling it in.
-
-But I already had this written and haven't yet re-done it that way. Feel free to redo and send a pull request on GitHub :)
-
-Again, it is largely a matter of preference and what makes the most sense to decide if one wants to have code inside Quarto files, or as separate R files.
-And sometimes, an R script with enough comments is good enough and one doesn't need a Quarto file.
-
-Also note that while here I split cleaning and exploring, this is iterative.
-You saw that as part of the processing, we already had to explore the data somewhat to understand how to clean it.
-In general, as you explore, you'll find things that need cleaning.
-As you clean, you can explore more.
-Therefore, at times it might make more sense to combine the cleaning and exploring code parts into a single R or Quarto file.
-Or split things in any other logical way.
-
-As part of the exploratory analysis, you should produce plots or tables or other summary quantities for the most interesting/important quantities in your data.
-Depending on the total number of variables in your dataset, explore all or some of the others.
-Figures produced here might be histograms or density plots, correlation plots, etc.
-Tables might summarize your data.
-
-Start by exploring one variable at a time.
-Then continue by creating plots or tables of the outcome(s) of interest and the predictor/exposure/input variables you are most interested in.
-If your dataset is small, you can do that for all variables.
-
-Plots produced here can be scatterplots, boxplots, violinplots, etc. Tables can be simple 2x2 tables or larger ones.
-
 # Setup
-
+Necessary libraries for this script.
 ```{r}
-#load needed packages. make sure they are installed.
-library(here) #for data loading/saving
+library(here) 
 library(dplyr)
@@
 ```{r}
-#Path to data. Note the use of the here() package and not absolute paths
 data_location <- here::here("data","processed-data","processeddata1.rds")
-#load data
 mydata <- readRDS(data_location)
@@
 
-Showing a bit of code to produce and save a summary table.
+Creating a summary table.
 
@@
 
-We are saving the results to the `results` folder.
-Depending on how many tables/figures you have, it might make sense to have separate folders for each.
-And/or you could have separate folders for exploratory tables/figures and for final tables/figures.
-Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.
-
 # Data exploration through figures
 
-Histogram plots for the continuous outcomes.
+Histogram plots for some continuous outcomes.
 
-Height first.
+Age first.
 
@@
 
-Now weights.
+Next bank account balance.
 
@@
 
-Now height as function of weight.
+Now balance as function of age
 
@@
 
-Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.
+Now balance as function of age, stratified by the marital variable.
 
@@
 ```
+
+Next, exploring balance across different job types.
 
@@
 
+Now, some barplots examining job types, education level, and marital status.
+
 ```{r}
-
 mydata |>
@@
   geom_bar()
-
 ```
-
-# Notes
-
-For your own explorations, tables and figures can be "quick and dirty".
-As long as you can see what's going on, there is no need to polish them.
-That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.).
-Those should look as nice, polished and easy to understand as possible.

### 2024-06-14 code/processing-code/processingfile-v1 - my version.qmd
no more template info

@@
 ---
-title: "An example cleaning script"
+title: "Preparing data for analysis"
 author: "Antonio Flores"
@@
 
-# Processing script
-
-This contains the same code and comments/information as `processingcode.R`.
-
-This just shows it as Quarto file, to give you an idea how to do it in a setup that combines code and text in a single file.
-
-See the other Quarto file for my currently preferred approach of pulling code from the R script into the Quarto file.
-
 # Setup
 
-Load needed packages. make sure they are installed.
+The following packages are needed for this data processing
 
@@
 
-Note that for functions that come from specific packages (instead of base R), I often specify both package and function like so:
-package::function() that's not required one could just call the function specifying the package makes it clearer where the function "lives",
-but it adds typing. You can do it either way.
-
 ```{r}
-# path to data
-# note the use of the here() package and not absolute paths
-
 data_location <- here::here("data","raw-data","bank-full.csv")
@@
 
-First we can look at the codebook
+Examning metadata by accessing the codebook
 
@@
 
-Several ways of looking at the data
+Various ways of looking at the data
 
@@
 
-# Cleaning
+# Cleaning the data
+
+We are mainly converting character variables into factors so we can do some categorical analysis
 ```{r}
-
 d1 = rawdata |> 
@@
 summary(d1)
-
 ```
@@
 
-Let's assign at the end to some final variable, this makes it easier to add further cleaning steps above.
+Assigning data to output variable.
 
@@
 
-Finally, we save the clean data as RDS file. I suggest you save your processed and cleaned data as RDS or RDA/Rdata files. 
-This preserves coding like factors, characters, numeric, etc.  If you save as CSV, that information would get lost.
-However, CSV is better for sharing with others since it's plain text. If you do CSV, you might want to write down somewhere what each variable is.
-
-See here for some suggestions on how to store your processed data:
-http://www.sthda.com/english/wiki/saving-data-into-r-data-format-rds-and-rdata
-
 ```{r}
@@
 ```
-
-# Notes
-
-Removing anyone who had "faulty" or missing data is one approach.
-It's often not the best.
-based on your question and your analysis approach, you might want to do cleaning differently (e.g.
-keep individuals with some missing information).

### 2024-06-14 products/manuscript/Manuscript.qmd
no more template info

@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
-# load a few R packages
 library(here)
@@
 
-_To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:_
-
-_Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]._
-
 {{< pagebreak >}}
@@
 I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already.
-Then I will test for correllation between different features and proceed from there.
+Then I will test for correlation between different features and proceed from there.
 
 ## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
 
@@
 
-raw = read.csv("C:/Users/Client/Documents/Flores-P2-Project/bank-full.csv",header = TRUE,
-                  sep=";")
-head(raw)
-
 ```{r}
@@
 #| echo: FALSE
-raw=read.csv("../../bank-full.csv", header = TRUE, sep=";")
+data_location <- here::here("data","raw-data","bank-full.csv")
+raw=read.csv(data_location, header = TRUE, sep=";")
 head(raw)
@@
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
 @tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
 
@@
 #| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
+resulttable=readRDS("../../results/tables/summarytable1.rds")
 knitr::kable(resulttable)
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+@fig-result shows a scatter plot figure produced by one of the R scripts.
 
@@
 #| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
+#| fig-cap: "Age and bank account balance stratified by marital status"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
+knitr::include_graphics(here("results","figures","age-balance-stratified.png"))
 ```
@@
 ## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
 
@@
 #| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
+#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
+#knitr::kable(resulttable2)
 ```
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
 
 ## Conclusions
-_What are the main take-home messages?_
-
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
 

### 2024-06-14 code/analysis-code/statistical-analysis - my version.R

@@ -1,19 +1,13 @@
 ###############################
-# analysis script
-#
-#this script loads the processed, cleaned data, does a simple analysis
-#and saves the results to the results folder
 
-#load needed packages. make sure they are installed.
+#required libraries
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
 
-#path to data
-#note the use of the here() package and not absolute paths
-data_location <- here::here("data","processed-data","processeddata.rds")
+data_location <- here::here("data","processed-data","processeddata1.rds")
 
-#load data. 
+#loading data. 
 mydata <- readRDS(data_location)
 
 

### 2024-06-14 code/eda-code/eda - my version.qmd
New plots, Updated Manuscript

@@
 library(scales)
+library(tidyr)
+library(forcats)
 ```
 
-Load the data.
+Loading the data.
 
@@
 ```{r}
-mydata |>
-  ggplot(aes(job))+
-  geom_bar()
+p6 = mydata |>
+  ggplot(aes(x=fct_infreq(job)))+
+  geom_bar() +
+  coord_flip()
 
-mydata |>
+p7 = mydata |>
   ggplot(aes(education))+
@@
 
-mydata |>
+p8 = mydata |>
   ggplot(aes(marital))+
   geom_bar()
+
+p6
+p7
+p8
+
+figure_file = here("results", "figures", "job-barchart.png")
+ggsave(filename = figure_file, plot=p6) 
+figure_file = here("results", "figures", "education-barchart.png")
+ggsave(filename = figure_file, plot=p7) 
+figure_file = here("results", "figures", "marital-barchart.png")
+ggsave(filename = figure_file, plot=p8) 
 ```
+
+Quick look at positive cases vs negative cases of outcome variable (y_termSubscribed)
+```{r}
+plot(mydata$y_termSubscribed)
+```
+
+Lets see how the age demographic differs between the population that subscribed to a term (y_termsubscribed=yes) vs the general population.
+
+```{r}
+p9 = mydata |>
+  filter(y_termSubscribed == "yes") |>
+  ggplot(aes(age))+
+  geom_bar()+
+  labs(title = "yes")
+
+p10 = mydata |>
+  ggplot(aes(age))+
+  geom_bar()+
+  labs(title = "yes/no")
+
+p9
+p10
+```
+
+Now, lets assess the highest job categories that are labeled as subscribing to a term.
+
+```{r}
+p11 = mydata |>
+  filter(y_termSubscribed == "yes") |>
+  ggplot(aes(x=fct_infreq(job)))+
+  geom_bar()+
+  coord_flip()+
+  labs(title = "Term Subscribed = yes")
+
+p11
+```

### 2024-06-14 code/processing-code/processingfile-v1 - my version.qmd
New plots, Updated Manuscript

@@
 ```{r}
-library(readxl) #for loading Excel files
-library(dplyr) #for data processing/cleaning
-library(tidyr) #for data processing/cleaning
-library(skimr) #for nice visualization of data 
-library(here) #to set paths
+library(readxl) 
+library(dplyr) 
+library(tidyr) 
+library(skimr) 
+library(here) 
 ```
@@
 
-Examning metadata by accessing the codebook
+Examining metadata by accessing the codebook.
+This codebook was taken directly from the data repository housing the chosen dataset.
+The decision was made to keep this in a separate xlsx file from the dataset so that the dataset could remain in the identical form in which it was sourced (CSV format).
 
@@
 
-We are mainly converting character variables into factors so we can do some categorical analysis
+We are mainly converting character variables into factors so we can do some categorical analysis.
+I had initially included converting months to a numeric variable but realized that it was fine as a factor variable.
 ```{r}
@@
   mutate(y = as.factor(y))
+
+colnames(d1)[17] <- "y_termSubscribed"
 

### 2024-06-14 products/manuscript/Manuscript.qmd
New plots, Updated Manuscript

@@
 Additionally, the classification variable is whether or not the person subscribed to a term deposit.
-There are 16 features in total.
+There are 17 features in total.
+
+Among the variables are a handful of features relating to the marketing campaign itself.
+For example, included are the day, month, and duration of the call, the number of contacts performed during the campaign (campaign), the number of days since the client was last contacted (pdays), the number of contacts performed before this campaign (previous), and finally the outcome of precious marketing campaigns (poutcome).
 
@@
 
-@tbl-summarytable shows a summary of the data.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable1.rds")
-knitr::kable(resulttable)
-```
-
-## Basic statistical analysis
-
-@fig-result shows a scatter plot figure produced by one of the R scripts.
+@fig-result shows a boxplot figure comparing balance levels across job types.
 
@@
 #| label: fig-result
+#| fig-cap: "Job type and bank account balance"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","job-balance-scatter.png"))
+```
+
+@fig-result1 shows a bar chart showing education levels of the data.
+
+```{r}
+#| label: fig-result1
+#| fig-cap: "Education level"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","education-barchart.png"))
+```
+
+@fig-result2 shows a scatter plot figure produced by one of the R scripts.
+
+```{r}
+#| label: fig-result2
 #| fig-cap: "Age and bank account balance stratified by marital status"
@@
 
-## Full analysis
+## Basic statistical analysis
 
@@
 ```
+
+## Full analysis
 

### 2024-06-14 code/analysis-code/statistical-analysis - my version.R

@@ -17,34 +17,16 @@ mydata <- readRDS(data_location)
 
 ############################
 #### First model fit
-# fit linear model using height as outcome, weight as predictor
+# fit linear model using balance as outcome, age as predictor
 
-lmfit1 <- lm(Height ~ Weight, mydata)  
+lmfit1 <- lm(balance ~ age, mydata)  
 
 # place results from fit into a data frame with the tidy function
 lmtable1 <- broom::tidy(lmfit1)
 
 #look at fit results
 print(lmtable1)
-
+summary(lmfit1)
 # save fit results table  
 table_file1 = here("results", "tables", "resulttable1.rds")
 saveRDS(lmtable1, file = table_file1)
-
-############################
-#### Second model fit
-# fit linear model using height as outcome, weight and gender as predictor
-
-lmfit2 <- lm(Height ~ Weight + Gender, mydata)  
-
-# place results from fit into a data frame with the tidy function
-lmtable2 <- broom::tidy(lmfit2)
-
-#look at fit results
-print(lmtable2)
-
-# save fit results table  
-table_file2 = here("results", "tables", "resulttable2.rds")
-saveRDS(lmtable2, file = table_file2)
-
-  

### 2024-06-22 code/eda-code/eda_my version.qmd
Part 3

--- 2024-06-14 code/eda-code/eda - my version.qmd
@@
 title: "Exploratory Data Analysis script"
-date: "2024-02-07"
+date: "2024-06-20"
 output: html_document
@@
 ```{r}
+#| echo: false
+#| output: false
 library(here) 
@@
 library(forcats)
+library(corrplot)
 ```
@@
 ```{r}
+#If this doesn't initially work, try restarting R and then proceeding
 data_location <- here::here("data","processed-data","processeddata1.rds")
@@
 
-Quick look at positive cases vs negative cases of outcome variable (y_termSubscribed)
-```{r}
-plot(mydata$y_termSubscribed)
-```
-
 Lets see how the age demographic differs between the population that subscribed to a term (y_termsubscribed=yes) vs the general population.
@@
 p9 = mydata |>
-  filter(y_termSubscribed == "yes") |>
+  filter(y_termSubscribed == 1) |>
   ggplot(aes(age))+
@@
 p10
+
+#Comparing how age daa compares from those who subscribed to a term deposit and those
+#that did not
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=age))+
+  facet_wrap(~y_termSubscribed, nrow=2)
+
 ```
@@
 p11 = mydata |>
-  filter(y_termSubscribed == "yes") |>
+  filter(y_termSubscribed == 1) |>
   ggplot(aes(x=fct_infreq(job)))+
@@
 ```
+
+Trying to explore some additional aspects of the data
+
+This shows the relationship between age and balance, grouped by education
+```{r}
+ggplot(data=mydata)+
+  geom_point(mapping =aes(x=age, y=balance))+
+  facet_wrap(~education, nrow=2)
+
+```
+
+Next, we can assess how age relates the number of jobs, grouped by job type.
+```{r}
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=age))+
+  facet_wrap(~job, nrow=2)
+```
+
+Digging into the Duration variable, we can see some very extreme values affecting this variable
+```{r}
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=duration))
+
+boxplot(mydata$duration)
+
+```
+
+Taking a quick look at some of the binary variables in the dataset
+```{r}
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=default))
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=housing))
+ggplot(data=mydata)+
+  geom_bar(mapping =aes(x=loan))
+```
+
+Taking a look at frequency counts across different variables of interest
+```{r}
+hist(numericvar$age)
+hist(numericvar$balance)
+hist(numericvar$duration)
+hist(numericvar$campaign)
+hist(numericvar$pdays)
+hist(numericvar$previous)
+```
+
+Examining which days are most common for recording a positive outcome (subscribing to 
+a term deposit). Looks like later in the month has higher frequencies.
+
+```{r}
+p12= mydata |>
+  filter(y_termSubscribed == 1) |>
+  ggplot(aes(day))+
+  geom_bar()+
+  labs(title = "yes")
+
+p12
+```
+
+We've analyzed which job types had the MOST positive outcomes, but now we want to 
+compare which job type had the greates PERCENT of positive outcomes.
+
+```{r}
+posjob = mydata %>% 
+  filter(y_termSubscribed==1) %>% 
+  count(y_termSubscribed, job, sort=TRUE)
+
+jobtable = mydata %>% count(job)
+
+pctjob = posjob %>% 
+  left_join(jobtable, by ="job")
+
+colnames(pctjob)[3] <- "numofYES"
+colnames(pctjob)[4] <- "numofBOTH"
+
+ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
+  geom_bar(stat="identity")+
+  coord_flip()
+
+ggplot(pctjob, aes(x=reorder(job, (numofYES/numofBOTH)), y=(numofYES/numofBOTH)))+
+  geom_bar(stat="identity")+
+  coord_flip()
+
+```
+On the first figure, we can see that jobs with the most positive outcomes are management, technician, and blue-collar, but on the second chart we can see that the job-types with the
+highest percentages are 'student', 'retired', and 'unemployed.' There appears to be a noticable different between the total number of positive cases, and the job types which are most commonly recording positive cases.

### 2024-06-22 code/processing-code/processingfile-v1 - my version.qmd
Part 3

@@
 ```{r}
+#| echo: false
+#| output: false
 library(readxl) 
@@
   mutate(poutcome = as.factor(poutcome)) |>
-  mutate(y = as.factor(y))
+  mutate(day = as.factor(day)) |>
+  mutate(y = ifelse(y == "yes", 1, 0))
 

### 2024-06-22 products/manuscript/Manuscript.qmd
Part 3

@@
 This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.
+After the data has been cleaned and prepared for analysis, exploratory data analysis will be performed to gather more information regarding the shape, size, and behavior of different variables, to gauge their usefulness in a prediction model.
+Additionally, several different statistical tests will be implemented to identify which classes within different variables affect the outcome variable.
+Finally (_haven't started on this yet_), with some additional data preprocessing, this project will test different machine learning models and conclude with the resulting findings.
 
@@
 ## General Background Information
-Providing businesses with a model that will allow prioritization of consumers could save time and allow for a more accurate use of resources.
+Providing businesses with a model that will allow prioritization of consumers and/or demographics has great potential in improving resource management and future marketing campaigns, as well as increasing efficient spending.
 
@@
 
-The research question I plan to address with my analysis is which features or combination of features are the best predictors of consumers making a deposit.
+The research question I plan to address with my analysis is: which features or combination of features are the best predictors of consumers making a deposit?
 The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns.
-I don't currently have specific predictors in mind to begin with, as I plan to test all of the available features while also looking at some combinations or concatenation of features.
+Currently, I plan to investigate all variables, but I am specifically interested in both job type, education and age.
 
@@
 
-I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already.
-Then I will test for correlation between different features and proceed from there.
+Cleaning -> This includes converting numeric variables to factors, changing column names, and converting the positive outcome varible to 1 and 0.
 
-## Schematic of workflow
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-```
+Exploring -> Using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.
 
 ## Data aquisition
-Data retrieved from UCI ML Repository
+The dataset for this project was retrieved from UCI ML Repository in CSV form.
+Additionally, I created a codebook based on data from the same source.
 
@@
 
-## Statistical analysis
-
 {{< pagebreak >}}
@@
 
+@fig-result4 shows barplot of age data for the full dataset
+
+```{r}
+#| label: fig-result4
+#| fig-cap: "Age data for those with positive classifcation"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","yes-no-age.png"))
+```
+
+@fig-result3 shows barplot of age data for only those who subscribed to a term deposit
+
+```{r}
+#| label: fig-result3
+#| fig-cap: "Age data for those with positive classifcation"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","yes-age.png"))
+```
+
+@fig-result5 shows barplot of most common days of the month to record a positive outcome.
+
+```{r}
+#| label: fig-result5
+#| fig-cap: "Days of the month for positive outcomes"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","days-yes.png"))
+```
+
+@fig-result6 Job type for count of positive outcomes.
+
+```{r}
+#| label: fig-result6
+#| fig-cap: "Job type for count of positive outcomes."
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","jobs-count-yes.png"))
+```
+
+@fig-result7 Job type for percent of positive outcomes per job type.
+
+```{r}
+#| label: fig-result7
+#| fig-cap: "Job type for percent of positive outcomes per job type."
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","jobs-percent-yes.png"))
+```
+
 ## Basic statistical analysis
@@
 
+#WORK IN PROGRESS
+
 ## Full analysis
+
+_Pending Data Preprocessing and ML Modeling_
+_I plan to use recipies and some of the tinymodels for this section_
 

### 2024-06-22 code/analysis-code/statistical-analysis - my version.R

@@ -4,6 +4,7 @@
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
+library(jmv) # for performing MANCOVA
 
 data_location <- here::here("data","processed-data","processeddata1.rds")
 
@@ -19,7 +20,7 @@ mydata <- readRDS(data_location)
 #### First model fit
 # fit linear model using balance as outcome, age as predictor
 
-lmfit1 <- lm(balance ~ age, mydata)  
+lmfit1 <- lm(balance ~ y_termSubscribed, mydata)  
 
 # place results from fit into a data frame with the tidy function
 lmtable1 <- broom::tidy(lmfit1)
@@ -30,3 +31,66 @@ summary(lmfit1)
 # save fit results table  
 table_file1 = here("results", "tables", "resulttable1.rds")
 saveRDS(lmtable1, file = table_file1)
+
+
+#Fitting several different linear models to explore the relationships among variables
+
+lm(y_termSubscribed ~ education, data = mydata) %>% 
+  summary()
+#Significant, R^2: .5%
+
+lm(y_termSubscribed ~ job, data = mydata) %>% 
+  summary()
+#Significant, R^2: 1.8%
+
+lm(y_termSubscribed ~ day, data = mydata) %>% 
+  summary()
+#Significant, R^2: 1.2%
+
+lm(y_termSubscribed ~ age, data = mydata) %>% 
+  summary()
+#Significant, R^2: .06%
+
+lm(y_termSubscribed ~ month, data = mydata) %>% 
+  summary()
+#Significant, R^2: 6.75%
+#Largest R^2 so far!
+
+lm(y_termSubscribed ~ duration, data = mydata) %>% 
+  summary()
+#Significant, R^2 is at 15%
+#Duration was noted as a predictor that could lead to overfitting. 
+#We will most likely not use this one
+
+
+#Trying Multi-Variable Linear Models
+
+lm(y_termSubscribed ~ age + job + education + day + month, data = mydata) %>% 
+  summary()
+#Significant, R^2: 8.6%
+
+
+### Logistic Regression
+mylogit = glm(y_termSubscribed ~ job + education, data=mydata, family="binomial")
+summary(mylogit)
+
+#Factors that stand out: For jobs, it looks like 'retired', 'student', and 'unemployed' lead to 
+#greater odds of subscribing to a deposit.
[…]

### 2024-06-22 code/eda-code/eda_my version.qmd
minor edit

@@
 
-#Comparing how age daa compares from those who subscribed to a term deposit and those
+figure_file = here("results", "figures", "yes-age.png")
+ggsave(filename = figure_file, plot=p9)
+
+figure_file = here("results", "figures", "yes-no-age.png")
+ggsave(filename = figure_file, plot=p10)
+
+#Comparing how age data compares from those who subscribed to a term deposit and those
 #that did not
@@
 p12
+figure_file = here("results", "figures", "days-yes.png")
+ggsave(filename = figure_file, plot=p12) 
 ```
@@
 
-ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
+p13 = ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
   geom_bar(stat="identity")+
@@
 
-ggplot(pctjob, aes(x=reorder(job, (numofYES/numofBOTH)), y=(numofYES/numofBOTH)))+
+p14 = ggplot(pctjob, aes(x=reorder(job, (numofYES/numofBOTH)), y=(numofYES/numofBOTH), fill=numofYES))+
   geom_bar(stat="identity")+
@@
 
+figure_file = here("results", "figures", "jobs-count-yes.png")
+ggsave(filename = figure_file, plot=p13) 
+figure_file = here("results", "figures", "jobs-percent-yes.png")
+ggsave(filename = figure_file, plot=p14)
+
+p13
+p14
 ```

### 2024-06-22 products/manuscript/Manuscript.qmd
minor edit

@@
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+Example @tbl-resulttable4 shows a summary of a linear model fit.
 
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: tbl-resulttable4
+#| tbl-cap: "Logistic regression model fit table."
 #| echo: FALSE
-#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-#knitr::kable(resulttable2)
+resulttable4 = readRDS(here("results","tables","resulttable4.rds"))
+knitr::kable(resulttable4)
 ```

### 2024-06-22 code/analysis-code/statistical-analysis - my version.R

@@ -74,6 +74,11 @@ lm(y_termSubscribed ~ age + job + education + day + month, data = mydata) %>%
 mylogit = glm(y_termSubscribed ~ job + education, data=mydata, family="binomial")
 summary(mylogit)
 
+logittable1 <- broom::tidy(mylogit)
+
+table_file2 = here("results", "tables", "resulttable4.rds")
+saveRDS(logittable1, file = table_file2)
+
 #Factors that stand out: For jobs, it looks like 'retired', 'student', and 'unemployed' lead to 
 #greater odds of subscribing to a deposit.
 #For education, 'secondary', 'tertiary', and 'unknown' all have better odds than 'primary'

### 2024-07-13 code/eda-code/eda_my version.qmd
MAJOR CHANGES

@@
 library(corrplot)
+library(corrplot)
 ```
@@
 print(summary_df)
+
 # save to file
@@
 saveRDS(summary_df, file = summarytable_file)
+```
+
+```{r}
+library(knitr)
+library(kableExtra)
+
+data_location <- here::here("data","processed-data","processeddata1.rds")
+mydata1 <- readRDS(data_location)
+
+mydata1 = mydata1 %>% 
+  select(age, job, marital, education, default, balance, housing, loan, contact, day, month, y_termSubscribed)
+
+str_summary <- capture.output(str(mydata1))
+
+describetbl1 = kable(str_summary, caption = "Summary of predictors") %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
+
+summarytable_file = here("results", "tables", "describetbl1.rds")
+saveRDS(describetbl1, file = summarytable_file)
+
+describetbl1
 ```
@@
 
-Trying to explore some additional aspects of the data
-
-This shows the relationship between age and balance, grouped by education
-```{r}
-ggplot(data=mydata)+
-  geom_point(mapping =aes(x=age, y=balance))+
-  facet_wrap(~education, nrow=2)
-
-```
-
-Next, we can assess how age relates the number of jobs, grouped by job type.
-```{r}
-ggplot(data=mydata)+
-  geom_bar(mapping =aes(x=age))+
-  facet_wrap(~job, nrow=2)
-```
-
-Digging into the Duration variable, we can see some very extreme values affecting this variable
-```{r}
-ggplot(data=mydata)+
-  geom_bar(mapping =aes(x=duration))
-
-boxplot(mydata$duration)
-
-```
-
 Taking a quick look at some of the binary variables in the dataset
@@
 ```{r}
-hist(numericvar$age)
-hist(numericvar$balance)
-hist(numericvar$duration)
-hist(numericvar$campaign)
-hist(numericvar$pdays)
-hist(numericvar$previous)
+hist(mydata$age)
+hist(mydata$balance)
 ```
@@
 We've analyzed which job types had the MOST positive outcomes, but now we want to 
-compare which job type had the greates PERCENT of positive outcomes.
+compare which job type had the greatest PERCENT of positive outcomes.
 
@@
 
+```
+
+```{r}
 p13 = ggplot(pctjob, aes(x=reorder(job, numofYES), y=numofYES, fill=numofYES))+
@@
 highest percentages are 'student', 'retired', and 'unemployed.' There appears to be a noticable different between the total number of positive cases, and the job types which are most commonly recording positive cases.
+
+```{r}
+data_location <- here::here("data","processed-data","processeddata5.rds")
+mydata <- readRDS(data_location)
+
+numericonly = mydata %>% 
+  select(!y_termSubscribed)
+
+# Specify the file path where the image will be saved
+file_path <- here("results", "figures","corplot.png")
+
+# Open a PNG graphics device
+png(filename = file_path, width = 800, height = 600)
+
+# Create the correlation plot
+corrplot(cor(numericonly))
+
+# Close the graphics device
+dev.off()
+```

### 2024-07-13 code/processing-code/processingfile-v1 - my version.qmd
MAJOR CHANGES

@@
 I had initially included converting months to a numeric variable but realized that it was fine as a factor variable.
+
 ```{r}
@@
   mutate(month = as.factor(month)) |>
+  mutate(y = ifelse(y == "yes", 1, 0)) |>
+  mutate(y = as.factor(y)) |>
   mutate(poutcome = as.factor(poutcome)) |>
   mutate(day = as.factor(day)) |>
-  mutate(y = ifelse(y == "yes", 1, 0))
+  mutate(y = as.factor(y)) %>% 
+  select(!duration)
 
-colnames(d1)[17] <- "y_termSubscribed"
+colnames(d1)[16] <- "y_termSubscribed"
 
@@
 ```
+
+# Attempting to convert categorical variables to dummy
+
+```{r}
+library(caret)
+
+testing_dummy = processeddata1 %>% 
+  select(job, marital, education, default, housing, loan, contact) %>% 
+  mutate(ID = row_number())
+
+dummy <- dummyVars("~.", data=testing_dummy)
+df1 <- predict(dummy, newdata = testing_dummy)
+
+df1 = as.data.frame(df1)
+
+numeric_vars = processeddata1 %>% 
+  select(age, balance, day, y_termSubscribed) %>% 
+  mutate(ID = row_number())
+
+d2 <- merge(df1,numeric_vars, by = "ID")
+
+```
+
+```{r}
+save_data_location <- here::here("data","processed-data","processeddata2.rds")
+saveRDS(d2, file = save_data_location)
+```
+
+# Attempting to convert categorical variables to numeric factors
+
+```{r}
+library(caret)
+
+d3 = processeddata1 |>
+  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
+  mutate(job = as.factor(unclass(job))) |>
+  mutate(marital = as.factor(unclass(marital))) |>
+  mutate(education = as.factor(unclass(education))) |>
+  mutate(default = as.factor(unclass(default))) |>
+  mutate(housing = as.factor(unclass(housing))) |>
+  mutate(loan = as.factor(unclass(loan)))
+
+summary(d3$education)
+```
+
+```{r}
+save_data_location <- here::here("data","processed-data","processeddata3.rds")
+saveRDS(d3, file = save_data_location)
+```
+
+# Attempting to convert categorical variables to numeric
+
+```{r}
+library(caret)
+
+d4 = processeddata1 |>
+  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
+  mutate(job = unclass(job)) |>
+  mutate(marital = unclass(marital)) |>
+  mutate(education = unclass(education)) |>
+  mutate(default = unclass(default)) |>
+  mutate(housing = unclass(housing)) |>
+  mutate(loan = unclass(loan))|>
+  mutate(day = unclass(day))
+
+```
+
+```{r}
+save_data_location <- here::here("data","processed-data","processeddata4.rds")
+saveRDS(d4, file = save_data_location)
+```
+
+# Attempting to convert categorical variables to true numeric
+
+```{r}
+library(caret)
+
+d5 = processeddata1 |>
+  select(job, marital, education, default, housing, loan, age, balance, day, y_termSubscribed) |>
+  mutate(job = as.numeric(job)) |>
+  mutate(marital = as.numeric(marital)) |>
+  mutate(education = as.numeric(education)) |>
+  mutate(default = as.numeric(default)) |>
+  mutate(housing = as.numeric(housing)) |>
+  mutate(loan = as.numeric(loan))|>
+  mutate(day = as.numeric(day))
+
+```
+
+```{r}
+save_data_location <- here::here("data","processed-data","processeddata5.rds")
+saveRDS(d5, file = save_data_location)
+```

### 2024-07-13 products/manuscript/Manuscript.qmd
MAJOR CHANGES

--- 2024-06-14 products/manuscript/Manuscript.qmd
@@
 ---
-title: "Working Title: Predicting Customer Behavior Using a Portuguese Financial Institution Dataset "
+title: "Predicting Customer Behavior Using a Portuguese Financial Institution Dataset "
 subtitle: ""
@@
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning = FALSE}
 library(here)
@@
 # Summary/Abstract
+
 This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.
+After the data has been cleaned and prepared for analysis, exploratory data analysis will be performed to gather more information regarding the shape, size, and behavior of different variables, to gauge their usefulness in a prediction model.
+Additionally, several different statistical tests will be implemented to identify which classes within different variables affect the outcome variable.
+Finally this project will test different machine learning models and conclude with the resulting findings.
 
@@
 ## General Background Information
-Providing businesses with a model that will allow prioritization of consumers could save time and allow for a more accurate use of resources.
+
+Providing businesses with a model that will allow prioritization of consumers and/or demographics has great potential in improving resource management and future marketing campaigns, as well as increasing efficient spending.
+This particular project will utilize Classification Prediction.
+Unlike Regression Prediction which will attempt to predict a continuous value (e.g., x amount of dollars, x amount of cells), Classification Prediction seeks to train a model that can correctly predict a classification (e.g., True/False, Success/Failure) based on the given predictors.
+In our case, our y variable or response variable is a binary variable, Yes/No, answering whether or not a customer subscribed to a term deposit.
 
 ## Description of data and data source
+
 The data was donated on 2/13/2012.
@@
 
-There are 45,212 records, and includes: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history.
-Additionally, the classification variable is whether or not the person subscribed to a term deposit.
-There are 17 features in total.
+There are 45,212 records,17 columns/variables which include: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history.
+Additionally, as mentioned, the classification variable is whether or not the person subscribed to a term deposit.
 
@@
 
-I want to first perform some EDA to identify if there are any variables with extreme variation or conversely if there some patterns that can be seen already.
-Then I will test for correlation between different features and proceed from there.
+## Data aquisition
 
-## Schematic of workflow
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-```
-
-## Data aquisition
-Data retrieved from UCI ML Repository
+The dataset for this project was retrieved from UCI ML Repository in CSV form. Additionally, I created a codebook based on data from the same source.
 
@@
 raw=read.csv(data_location, header = TRUE, sep=";")
-head(raw)
+kable(head(raw[1:5]))
+
+kable(head(raw[6:10]))
 ```
@@
 ### Dimensions:
-dim(raw)
 
-```{r}
-#| label: tbl-summarytable2
-#| tbl-cap: "Data Dimensions"
-#| echo: FALSE
-dim(raw)
+Rows: 45, 211\
+Columns: 17
+
+### Describing data
+
+```{r, echo=FALSE, message=FALSE, warning = FALSE}
+data_location <- here::here("results","tables","describetbl1.rds")
+describe1 <- readRDS(data_location)
+describe1
 ```
 
-### Describing raw data
-str(raw)
+### Cleaning Data
 
-```{r}
-#| label: tbl-summarytable3
-#| tbl-cap: "Data Description"
-#| echo: FALSE
-str(raw)
-```
+There were several different data cleaning methods that were experimented with in order to find the best way to prepare the data for modeling.
+First, I converted several variables to factors as they had been read in as character variables, or strings.
+Next I tried to use the DummyVars tool to convert every categorical variable to a dummy value but this made modeling difficult due to a large amount of binary variables.
+Instead I proceeded with converting the categorical variables to numeric values while maintaining their factor status.
+I also created a dataset with these numeric values that were not factors.
+These were the two primary dataset I used for my modeling.
+Finally I created a dataset stripping the categorical values of their attributes, this was solely created to allow for my corrplot to work.
 
-## Statistical analysis
+### Removing Initial Predictors
+
+I removed the 'Duration' variable due to a warning I discovered on the site hosting this dataset initially: "Important note: this attribute highly affects the output target (e.g., if duration=0 then y='no').
+Yet, the duration is not known before a call is performed.
+Also, after the end of the call y is obviously known.
+Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model."(@misc_bank_marketing_222) Due to this new information I decided to remove all the variables related to this, which included 'campaign', 'pdays', 'previous', and 'poutcome.' Bringing our new total number of predictors to 11.
 
@@
 
-@fig-result shows a boxplot figure comparing balance levels across job types.
+In this section we will be exploring the data using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.
 
-```{r}
-#| label: fig-result
-#| fig-cap: "Job type and bank account balance"
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","job-balance-scatter.png"))
-```
-
-@fig-result1 shows a bar chart showing education levels of the data.
+@fig-result1 shows the distribution of the 'Age' variable.
 
@@
 #| label: fig-result1
-#| fig-cap: "Education level"
+#| fig-cap: "Age Histogram"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","age-distribution.png"))
+```
+While this shows a slight skew to the left, if we account for the extreme value on the right, this data seems to be fairly normally distributed.
+
+@fig-result2 shows the distribution of the 'Education' variable
+
+```{r}
+#| label: fig-result2
+#| fig-cap: "Education level Bar Chart"
 #| echo: FALSE
@@
 
-@fig-result2 shows a scatter plot figure produced by one of the R scripts.
+@fig-result3 shows the most common 'Job' types in our data
 
 ```{r}
-#| label: fig-result2
+#| label: fig-result3
+#| fig-cap: "Most Common Job Types"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","job-barchart.png"))
+```
+
+Interestingly, Blue-collar and Management are the two most common job types. This play an important role later in our analysis due to the percent of the data these two categories represent.
+
+@fig-result4 shows a scatter plot figure produced by one of the R scripts.
+
+```{r}
+#| label: fig-result4
 #| fig-cap: "Age and bank account balance stratified by marital status"
@@
 
+@fig-result5 shows a barplot of most common days of the month to record a positive outcome.
+
+```{r}
+#| label: fig-result5
+#| fig-cap: "Days of the month for positive outcomes"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","days-yes.png"))
+```
+While there doesn't appear to be a significant trend, we can clearly see that the 30th of the month stands out as being a day of interest, especially compared to other days of the months.
+
+@fig-result6 shows potential correlation between different variales.
+
+```{r}
+#| label: fig-result6
+#| fig-cap: "Correlation Matrix"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","corplot.png"))
+```
+
+The biggest takeaways from our Exploratory Data Analysis is that most variables are fairly normally distributed due to the size of the data, and most variable are not correlated to each other, with one exception ('Age' and 'Marital')
+
 ## Basic statistical analysis
 
-Example @tbl-resulttable2 shows a summary of a linear model fit.
+Before we begin with the machine learning analysis, I sought to test a few variables of interest ('Age', 'Job') in their significance of affecting the response variable.
+As mentioned previously, our response variable is a binary value (Yes/No), which means that we need to use a logistic regression model instead of a linear regression model.
+
+Below are the results for our first logistic model fit with Age as Predictor.
 
 ```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
+#| label: basicmodel1
+#| fig-cap: "Logistic Regression with Age as Predictor"
 #| echo: FALSE
-#resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-#knitr::kable(resulttable2)
+
+data_location <- here::here("results","tables","basicmodel1.rds")
+basic1 <- readRDS(data_location)
+kable(basic1)
 ```
 
+Below are the results for our second logistic model fit with Job as Predictor.
+
+```{r}
+#| label: basicmodel2
+#| fig-cap: "Logistic Regression with Job as Predictor"
+#| echo: FALSE
+
+data_location <- here::here("results","tables","basicmodel2.rds")
+basic2 <- readRDS(data_location)
+kable(basic2)
+```
+
+As expected, both predictors are very significant. We will proceed to the Machine Learning portion of this project.
+
 ## Full analysis
+So far, I have focused on the MARS model, KNN, and Logistic Regression to determine the prediction model for this project.
+As highlighted in the data cleaning section, there were two datasets I was focusing on.
+One that had numeric factors for the categorical predictors, and one that had just numeric variables for the categorical predictors.
+I decided to examine both so there are a total of 6 different runs recorded below, 3 runs for each dataset.
+
+@tbl-resulttable1 displays the Accuracy and Kappa scores for the first three model runs using the predictors that are numeric and are not factors.
+```{r}
+#| label: tbl-resulttable1
+#| tbl-cap: "Performance Results for Models using Numeric (unfactored) Predictors"
+#| echo: FALSE
+resulttable1 = readRDS(here("results","tables","perftable1.rds"))
+knitr::kable(resulttable1)
+```
+
+@tbl-resulttable2 displays the Accuracy and Kappa scores for the first three model runs using the predictors that are numeric and are not factors.
+```{r}
+#| label: tbl-resulttable2
+#| tbl-cap: "Performance Results for Models using Numeric Factored Predictors"
+#| echo: FALSE
+resulttable2 = readRDS(here("results","tables","perftable2.rds"))
+knitr::kable(resulttable2)
+```
+
+In terms of accuracy, all 6 model runs seem to have almost identical accuracy. They do tend to differ quite a bit in terms of Kappa...
+
+Finally, we can examine which predictors specifically improved prediction the most.
+
+The following figures were taken from the Factored Numeric model runs.
+
+@fig-result7 shows the Predictors that had the most impact on the Mars Model
+
+```{r}
+#| label: fig-result7
+#| fig-cap: "Mars Model Variables of Importance"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","importantMM2.png"))
+```
+
+Housing2 indicates that the person does have a housing loan.\
+Loan2 indicates that the person does have a personal loan for an unspecified purpose.\
+Job9 indicates that the person is a Student.\
+education 3 indicates that the person has attained Tertiary Education.\
+
+@fig-result8 shows the Predictors that had the most impact on the KNN Model
+
+```{r}
+#| label: fig-result8
+#| fig-cap: "KNN Variables of Importance"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","importantKNN2.png"))
+```
 
@@
 ## Summary and Interpretation
-
+So far all of the models I have chosen have shown considerable predictive ability. I plan to tune the existing three models while also trying out Random Forest as well as potentially one more model. 
 ## Strengths and Limitations
@@
 ## Conclusions
+I will conclude with choosing the model that seems the most robust, accurate, and flexible, as well as determining which variables have the greatest predictive ability and which levels within those predictors stand out as particularly effective in classifying correctly.
 

### 2024-07-13 code/analysis-code/statistical-analysis - my version.R

@@ -5,97 +5,294 @@ library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
 library(jmv) # for performing MANCOVA
+library(earth)
+library(tidyr)
+library(dplyr)
+library(caret)
+library(knitr)
 
-data_location <- here::here("data","processed-data","processeddata1.rds")
+
+
+#loading data. 
+data_location <- here::here("data","processed-data","processeddata4.rds")
+mydata <- readRDS(data_location)
+summary(mydata)
+
+
+#Basic Statistical Analysis
+#Examining a couple logistic regression tests with two variables of interest
+
+
+logitAge = glm(y_termSubscribed~age, data = mydata, family = "binomial")
+summary(logitAge)
+
+lrtable1 <- broom::tidy(logitAge)
+basic_model1 = here("results", "tables", "basicmodel1.rds")
+saveRDS(lrtable1, file = basic_model1)
+
+logitJob = glm(y_termSubscribed~job, data=mydata, family = "binomial")
+summary(logitJob)
+
+lrtable2 <- broom::tidy(logitJob)
+basic_model2 = here("results", "tables", "basicmodel2.rds")
+saveRDS(lrtable2, file = basic_model2)
+
+#Looks like both of the above predictors are significant
+
+
+#Full Data Analysis
+
+
+######################################
+#Machine Learning Models Part 1
+######################################
+#Using Numeric Unfactored Variables
+############################
+
+
+#Setting Tran Control
+ctrl <- trainControl(method = "cv")
+
+#Splitting the data. 80% for training, 20% test.
+
+set.seed(21)
+split1 = sample(c(rep(0, 0.8 * nrow(mydata)), rep(1, 0.2*nrow(mydata))))
+
+
+train_data = mydata[split1 == 0, ]
+test_data = mydata[split1 == 1, ]
+
+train_y = train_data$y_termSubscribed
+test_y = test_data$y_termSubscribed
+
+
+train_x = train_data %>% 
+  select(!y_termSubscribed)
+
+test_x = test_data %>% 
+  select(!y_termSubscribed)
+
+
+#nearZeroVar(d2, saveMetrics = TRUE)
+
+#nzv(d2)
+
+
+
+
+#################
+#Modeling Section
+
+# MARS model
+set.seed(21)
+marsmodel = train(x= train_x, y=train_y, 
+                  method = "earth",
[…]

### 2024-07-28 products/manuscript/Manuscript.qmd
For Assn 5

@@
 
-# Summary/Abstract
+# Abstract
 
-This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese Bank.
-After the data has been cleaned and prepared for analysis, exploratory data analysis will be performed to gather more information regarding the shape, size, and behavior of different variables, to gauge their usefulness in a prediction model.
-Additionally, several different statistical tests will be implemented to identify which classes within different variables affect the outcome variable.
-Finally this project will test different machine learning models and conclude with the resulting findings.
+Financial institutions spend billions of dollars on their marketing teams every year, with a positive trend observed in 2023 for the largest and smallest banks (@akins2024).
+With the need for efficient marketing becoming more and more important, machine learning models are viable tools for refining marketing strategies.
+This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese bank by comparing the accuracy and predictive power of different machine learning models.
 
@@
 
+In 2022, JPMorgan Chase & Co.
+spent \$3.9 billion (US) dollars on Marketing, leading other financial institutions in this category.
+Recent reports have shown that this has only grown in past years (@jpmorgan2023).
 Providing businesses with a model that will allow prioritization of consumers and/or demographics has great potential in improving resource management and future marketing campaigns, as well as increasing efficient spending.
 This particular project will utilize Classification Prediction.
-Unlike Regression Prediction which will attempt to predict a continuous value (e.g., x amount of dollars, x amount of cells), Classification Prediction seeks to train a model that can correctly predict a classification (e.g., True/False, Success/Failure) based on the given predictors.
+Unlike Regression Prediction which will attempt to predict a continuous value (e.g., x amount of dollars, x amount of cells), Classification Prediction seeks to train a model that can correctly predict a classification (e.g., True/False, Success/Failure) based on the given predictors(@kuhn2018).
 In our case, our y variable or response variable is a binary variable, Yes/No, answering whether or not a customer subscribed to a term deposit.
@@
 
-There are 45,212 records,17 columns/variables which include: age, marital status, job, education,details related to the phone call, as well as answers related to questions about past credit history.
+There are 45,212 records,17 columns/variables which include: age, marital status, job, education, details related to the phone call, as well as answers related to questions about past credit history.
 Additionally, as mentioned, the classification variable is whether or not the person subscribed to a term deposit.
-
-Among the variables are a handful of features relating to the marketing campaign itself.
-For example, included are the day, month, and duration of the call, the number of contacts performed during the campaign (campaign), the number of days since the client was last contacted (pdays), the number of contacts performed before this campaign (previous), and finally the outcome of precious marketing campaigns (poutcome).
 
@@
 The desired output of this analysis is a model which allows a financial institution to better prioritize/make decisions regarding future marketing campaigns.
-Currently, I plan to investigate all variables, but I am specifically interested in both job type, education and age.
+I plan to investigate all demographic variables, with a specific focus on job type, education and age.
 
@@
 #| label: tbl-summarytable1
-#| tbl-cap: "Data Snapshot"
+#| tbl-cap: "Data Snapshot of first 5 Variables"
 #| echo: FALSE
@@
 kable(head(raw[1:5]))
+```
 
+```{r}
+#| label: tbl-summarytable2
+#| tbl-cap: "Data Snapshot of next 5 Variables"
+#| echo: FALSE
 kable(head(raw[6:10]))
+
 ```
@@
 Rows: 45, 211\
-Columns: 17
+Columns: 17 {{< pagebreak >}}
 
@@
 
-## Exploratory/Descriptive analysis
+## Exploratory Data Analysis
 
-In this section we will be exploring the data using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.
+Peng and Matsui define EDA as "the process of exploring your data","\[including\] examining the structure and components of your dataset, the distributions of individual variables, and the relationships between two or more variables"(@peng2018).
+
+In accordance with this definition, we will be exploring the data using different charts to identify outliers, abnormalities, relationships, and the general shape and feel of different variables.
 
@@
 #| label: fig-result1
-#| fig-cap: "Age Histogram"
 #| echo: FALSE
@@
 ```
-While this shows a slight skew to the left, if we account for the extreme value on the right, this data seems to be fairly normally distributed.
+
+While this shows a slight skew to the right, if we account for the extreme values, this data seems to be fairly normally distributed. {{< pagebreak >}}
 
@@
 
-Interestingly, Blue-collar and Management are the two most common job types. This play an important role later in our analysis due to the percent of the data these two categories represent.
+{{< pagebreak >}}
 
-@fig-result4 shows a scatter plot figure produced by one of the R scripts.
+@fig-result4 shows a scatter plot with Age and bank account balance as the (x,y) values, stratified by marital status.
 
@@
 #| label: fig-result4
-#| fig-cap: "Age and bank account balance stratified by marital status"
+#| fig-cap: "Scatterplot of Age and Balance"
 #| echo: FALSE
@@
 
-@fig-result5 shows a barplot of most common days of the month to record a positive outcome.
+@fig-result5 shows a barplot of the most common days of the month to record a positive outcome.
 
@@
 #| label: fig-result5
-#| fig-cap: "Days of the month for positive outcomes"
 #| echo: FALSE
@@
 ```
+
 While there doesn't appear to be a significant trend, we can clearly see that the 30th of the month stands out as being a day of interest, especially compared to other days of the months. 
+{{< pagebreak >}}
 
-@fig-result6 shows potential correlation between different variales.
+@fig-result6 shows potential correlation between different variables.
 
@@
 
-The biggest takeaways from our Exploratory Data Analysis is that most variables are fairly normally distributed due to the size of the data, and most variable are not correlated to each other, with one exception ('Age' and 'Marital')
+The biggest takeaways from our Exploratory Data Analysis is that most variables are fairly normally distributed due to the size of the data, and most variable are not correlated to each other, with one exception ('Age' and 'Marital') {{< pagebreak >}}
 
@@
 
-Before we begin with the machine learning analysis, I sought to test a few variables of interest ('Age', 'Job') in their significance of affecting the response variable.
-As mentioned previously, our response variable is a binary value (Yes/No), which means that we need to use a logistic regression model instead of a linear regression model.
+Before I began with the machine learning analysis, I sought to test the variables of interest in their significance of affecting the response variable.
+As mentioned previously, our response variable is a binary value (Yes/No), which means that we need to use a logistic regression model instead of a linear regression model(@kuhn2018).
 
-Below are the results for our first logistic model fit with Age as Predictor.
+Below are the results of the logistic model fit with all variables as predictors
 
@@
 #| label: basicmodel1
-#| fig-cap: "Logistic Regression with Age as Predictor"
 #| echo: FALSE
-
-data_location <- here::here("results","tables","basicmodel1.rds")
-basic1 <- readRDS(data_location)
-kable(basic1)
+data_location <- here::here("results","tables","fullmodel.rds")
+full1 <- readRDS(data_location)
+kable(full1)
 ```
 
-Below are the results for our second logistic model fit with Job as Predictor.
+As expected, all predictors are very significant as determined by the very small p-values. We will proceed to the Machine Learning portion of this project.
+
+## Machine Learning Modeling
+
+I chose to utilize the following models: the Multivariate Adaptive Regression Splines model (MARS), K-Nearest Neighbors model (KNN), Logistic Regression model, and RandomForest models to determine the best prediction model for this project.
+
+### Background on Chosen Models
+
+**MARS:**\
+The MARS model has some similarities to both Neural Networks and the Partial Least Squares model, but the distinguishing feature is a use of multiple 'splines' to create a "piecewise linear model" with multiple features modeling a separate part of the data (@kuhn2018).
 
 ```{r}
-#| label: basicmodel2
-#| fig-cap: "Logistic Regression with Job as Predictor"
+#| label: fig-result999
+#| fig-cap: "MARS Model example from MiniTab"
 #| echo: FALSE
+knitr::include_graphics(here("results","images","marsexample.jpeg"))
+```
+{{< pagebreak >}}
 
-data_location <- here::here("results","tables","basicmodel2.rds")
-basic2 <- readRDS(data_location)
-kable(basic2)
-```
+**KNN:**\
+The KNN model predicts based on the closest samples or neighbors.
+Essentially, to predict a value, the data is broken up into samples/neighbors, and then the nearest samples (using Euclidean distance, typically)to the value of interest are examined to either classify or find a mean between the chosen samples.
+K represents the number of neighbors to utilize to come to this conclusion (@kuhn2018).
 
-As expected, both predictors are very significant. We will proceed to the Machine Learning portion of this project.
+**Logistic Regression:**\
+As mentioned earlier, Logistic Regression is similar to Linear Regression but the difference is that a Logistic Regression focuses on the probability of an event (p, p-1).
 
-## Full analysis
-So far, I have focused on the MARS model, KNN, and Logistic Regression to determine the prediction model for this project.
-As highlighted in the data cleaning section, there were two datasets I was focusing on.
-One that had numeric factors for the categorical predictors, and one that had just numeric variables for the categorical predictors.
-I decided to examine both so there are a total of 6 different runs recorded below, 3 runs for each dataset.
+**RandomForest:**\
+RandomForest models take advantage of decision trees.
+If we think about the scenario in our project (whether someone makes a deposit or not), we could imagine a decsion tree starting with, "after what age is someone more likely to subscribe?" This would be our first node to split the data on.
+We could continue asking things like, "Are those with housing loans more likely to subscribe?" or "Are those who are married more liklely to subscribe?", and these would represent more decision nodes for us to split the data on, getting us closer to the mode accurate prediction model.
+The RandomForest algorithm uses different methods to create several uncorrelated "forests" of decision trees(@IBM) .
+{{< pagebreak >}}
+### Background Regarding Performance Metrics
 
-@tbl-resulttable1 displays the Accuracy and Kappa scores for the first three model runs using the predictors that are numeric and are not factors.
+As highlighted in the data cleaning section, there were two data sets I was focusing on: one that had numeric factors for the categorical predictors, and one that had just numeric variables for the categorical predictors.
+I decided to examine both so there are a total of 8 different runs recorded below, 4 runs for each dataset.
+
+**Description of Key Metrics:**\
+**Accuracy:**\
+The proportion of total correct classifications.\
+
+**Recall:**\
+The proportion of Positive cases that were correctly identified. This helps us understand how well we can classify positive cases specifically.\
+
+**Precision:**\
+The proportion of Positive classifications that were actually correct.\
+
+**F1:**\
+The harmonic mean of Precision and Recall. This metric helps iron out extreme values\
+{{< pagebreak >}}
+
+### Model Performance Metrics
+@tbl-resulttable1 displays the relevant metrics for the models runs using the predictors that are numeric and are not factors.
+
 ```{r}
 #| label: tbl-resulttable1
-#| tbl-cap: "Performance Results for Models using Numeric (unfactored) Predictors"
 #| echo: FALSE
-resulttable1 = readRDS(here("results","tables","perftable1.rds"))
+resulttable1 = readRDS(here("results","tables","finaltable1.rds"))
 knitr::kable(resulttable1)
@@
 
-@tbl-resulttable2 displays the Accuracy and Kappa scores for the first three model runs using the predictors that are numeric and are not factors.
+While all four models performed fairly similarly, the RandomForest model reported the highest values in all major categories.
+What stands out very clearly is that while all models have a fairly high accuracy scores, they also have very low recall values.
+We can deduce that this means the models are great at predicting negative cases (no subscription) but not very adequate at predicting positive cases (subscription).
+In the case of the Logistic Regression Model, there were 0 positive cases predicted at all, and thus the non-accuracy scores were 0 or NA.
+
+@tbl-resulttable2 displays the relevant metrics for the models using the predictors that are factors (Age/Balance are still numeric).
+
 ```{r}
 #| label: tbl-resulttable2
-#| tbl-cap: "Performance Results for Models using Numeric Factored Predictors"
 #| echo: FALSE
-resulttable2 = readRDS(here("results","tables","perftable2.rds"))
+resulttable2 = readRDS(here("results","tables","finaltable2.rds"))
 knitr::kable(resulttable2)
@@
 
-In terms of accuracy, all 6 model runs seem to have almost identical accuracy. They do tend to differ quite a bit in terms of Kappa...
+With these results we can see that the RandomForest Model is still the best performing, but in this case, the KNN model is not too far behind in most metrics.
+The Logistic Regression model performed better under these conditions, predicting some positive classes.
+However, the MARS model did not fare so well.
+In this run, the MARS model was the model to not identify a single positive class.
+{{< pagebreak >}}
+
+### Variables of Importance
 
 Finally, we can examine which predictors specifically improved prediction the most.
+That is, which had the greatest weight on the final result.
+We will compare the difference between the two different runs of the same models Note: I determined it would only be helpful to include models that had more an 0 positive classes predicted.
 
-The following figures were taken from the Factored Numeric model runs.
-
-@fig-result7 shows the Predictors that had the most impact on the Mars Model
+@fig-result7 shows the Predictors that had the most impact on the Mars Model -- Numeric
 
@@
 #| label: fig-result7
-#| fig-cap: "Mars Model Variables of Importance"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","importantMM2.png"))
+knitr::include_graphics(here("results","figures","importantM1.png"))
 ```
 
-Housing2 indicates that the person does have a housing loan.\
-Loan2 indicates that the person does have a personal loan for an unspecified purpose.\
-Job9 indicates that the person is a Student.\
-education 3 indicates that the person has attained Tertiary Education.\
+The numeric MARS models only highlighted a single variable of importance: Age.
+{{< pagebreak >}}
 
-@fig-result8 shows the Predictors that had the most impact on the KNN Model
+@fig-result9 shows the Predictors that had the most impact on the KNN Model--Numeric
 
 ```{r}
-#| label: fig-result8
+#| label: fig-result9
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","importantKNN1.png"))
+```
+
+Both KNN model runs returned the exact same most important factors, which were: Housing (whether someone had a housing loan or not), Balance (numeric value representing the customer's current account balance), and Education (Secondary, Tertiary, etc.).
+{{< pagebreak >}}
+@fig-result12 shows the Predictors that had the most impact on the Logistic Regression Model -- Factor
+
+```{r}
+#| label: fig-result12
 #| fig-cap: "KNN Variables of Importance"
 #| echo: FALSE
-knitr::include_graphics(here("results","figures","importantKNN2.png"))
+knitr::include_graphics(here("results","figures","importantLog2.png"))
 ```
+
+Housing2 indicates that the person does have a housing loan and Loan2 indicates that the person has personal loan.Education3 indicates that the person has attained a tertiary level of education.
+Additionally we see different days being flagged as important in determining accuracy.
+{{< pagebreak >}}
+@fig-result13 shows the Predictors that had the most impact on the RF Model -- Numeric
+
+```{r}
+#| label: fig-result13
+#| fig-cap: "RF Variables of Importance"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","importantRF1.png"))
+```
+{{< pagebreak >}}
+@fig-result14 shows the Predictors that had the most impact on the RF Model -- Factor
+
+```{r}
+#| label: fig-result14
+#| fig-cap: "KNN Variables of Importance"
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","importantRF2.png"))
+```
+
+For both RandomForest models, Age and Housing were the variables with most impact on the model's prediction power. Balance and Day were next for both models in different order of importance.
 
@@
 ## Summary and Interpretation
-So far all of the models I have chosen have shown considerable predictive ability. I plan to tune the existing three models while also trying out Random Forest as well as potentially one more model. 
-## Strengths and Limitations
+
+Initially the metrics produced by these models were showing substantial promise, but as I investigated the different confusion matrices I realized that the metrics were incorrectly labeling the negative classification as positive.
+When I reversed this option, the metrics very clearly shows that there was a drop in performance for predicting positive classifications.
+One explanation for this is that there were too few positive classes in the data as a whole.
+In the full data set, about 11% of the observations were positive cases, and this was also the case in the test data set as well.
+Given the size of the data set (45K records), and the small proportion of positive cases, there is reason to conclude that the models were victim to lack of familiarity with the positive cases.
+
+  Despite these findings, the models were able to produce which variables provided the most weight towards the prediction power.
+The Housing variable was listed as the most important variable by every model with the exception of the Mars-Numeric model.
+Balance and Age were also listed as important by several different models.
+The variables that did not show up in the five important variables in any of the models were: Marital, Job, and Default.
+Education, Loan, and Day had a significant effect on some of the models and little to no effect on others.
 
 ## Conclusions
-I will conclude with choosing the model that seems the most robust, accurate, and flexible, as well as determining which variables have the greatest predictive ability and which levels within those predictors stand out as particularly effective in classifying correctly.
+
+Using machine learning to better define a customer base can be an incredibly effective way to create more efficient marketing campaigns, as well as provide direction for better management of marketing resources.
+
+  I attempted to develop a classification model that accurately predicted whether a customer had subscribed to a term deposit or not.
+While the different models that were ran initially showed potential, the predictive power of the best model was not at an ideal level.
+With a more balanced data set with similar variables, it is probable that a more precise model could have been produced.
+
+  The other goal of this project was to identify which variables could best aid in identifying customers that would subscribe to a term deposit.
+I found that there was quite a bit of variance among the predictors in terms of importance, with some predictors clearly standing out above the rest.
+This information could be used to better segment a consumer pool or adjust marketing tactics to prioritize customers displaying the variables of importance uncovered in this project.
+
+  Further research would include utilizing additional models to confirm the results presented here, incorporating a more balanced data set (as mentioned above), and exploring different tuning options than were used in this project.
 

### 2024-07-28 code/analysis-code/statistical-analysis - my version.R

@@ -4,13 +4,13 @@
 library(ggplot2) #for plotting
 library(broom) #for cleaning up output from lm()
 library(here) #for data loading/saving
-library(jmv) # for performing MANCOVA
 library(earth)
 library(tidyr)
 library(dplyr)
 library(caret)
 library(knitr)
-
+library(randomForest)
+library(Metrics)
 
 
 #loading data. 
@@ -18,27 +18,17 @@ data_location <- here::here("data","processed-data","processeddata4.rds")
 mydata <- readRDS(data_location)
 summary(mydata)
 
-
 #Basic Statistical Analysis
-#Examining a couple logistic regression tests with two variables of interest
+#Examining a logistic regression test with all variables of interest
 
+logitFull = glm(y_termSubscribed~., data = mydata, family = "binomial")
+summary(logitFull)
 
-logitAge = glm(y_termSubscribed~age, data = mydata, family = "binomial")
-summary(logitAge)
+lrtable <- broom::tidy(logitFull)
+fullmodel = here("results", "tables", "fullmodel.rds")
+saveRDS(lrtable, file = fullmodel)
 
-lrtable1 <- broom::tidy(logitAge)
-basic_model1 = here("results", "tables", "basicmodel1.rds")
-saveRDS(lrtable1, file = basic_model1)
-
-logitJob = glm(y_termSubscribed~job, data=mydata, family = "binomial")
-summary(logitJob)
-
-lrtable2 <- broom::tidy(logitJob)
-basic_model2 = here("results", "tables", "basicmodel2.rds")
-saveRDS(lrtable2, file = basic_model2)
-
-#Looks like both of the above predictors are significant
-
+#All Predictors are Significant
 
 #Full Data Analysis
 
@@ -50,47 +40,48 @@ saveRDS(lrtable2, file = basic_model2)
 ############################
 
 
+#Preprocessing 2 True Numeric Variables
+set.seed(24)
+mydataPP = preProcess(mydata[7:8], method = c("scale", "center", "BoxCox"))
+mydata[7:8] = predict(mydataPP, mydata[7:8])
+
+
+#Splitting the data. 80% for training, 20% test.
+
+set.seed(21)
+split1 = sample(c(rep(0, 0.8 * nrow(mydata)), rep(1, 0.2*nrow(mydata))))
+
+
[…]

### 2024-08-05 manuscript/Manuscript.qmd
Last Major Changes

--- 2024-07-28 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/apa.csl
+bibliography: ../assets/dataanalysis-references.bib
+csl: ../assets/apa.csl
 ---
@@
 With the need for efficient marketing becoming more and more important, machine learning models are viable tools for refining marketing strategies.
-This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese bank by comparing the accuracy and predictive power of different machine learning models.
+This project will seek to identify a model for predicting consumer financial behavior using a dataset from a Portuguese bank.
+Through comparing the accuracy and predictive power of four different machine learning models, this project will identify an ideal method for prediction.
 
@@
 Rows: 45, 211\
-Columns: 17 {{< pagebreak >}}
+Columns: 17
+
+{{< pagebreak >}}
 
@@
 Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model."(@misc_bank_marketing_222) Due to this new information I decided to remove all the variables related to this, which included 'campaign', 'pdays', 'previous', and 'poutcome.' Bringing our new total number of predictors to 11.
-
-{{< pagebreak >}}
+#{{< pagebreak >}}
 
@@
 
-@fig-result2 shows the distribution of the 'Education' variable
-
 ```{r}
@@
 ```
-
-@fig-result3 shows the most common 'Job' types in our data
 
@@
 
-@fig-result4 shows a scatter plot with Age and bank account balance as the (x,y) values, stratified by marital status.
-
 ```{r}
 #| label: fig-result4
-#| fig-cap: "Scatterplot of Age and Balance"
+#| fig-cap: "Scatterplot of Age and Balance, stratified by Marital Status."
 #| echo: FALSE
@@
 While there doesn't appear to be a significant trend, we can clearly see that the 30th of the month stands out as being a day of interest, especially compared to other days of the months.
-{{< pagebreak >}}
 
@@
 
-The biggest takeaways from our Exploratory Data Analysis is that most variables are fairly normally distributed due to the size of the data, and most variable are not correlated to each other, with one exception ('Age' and 'Marital') {{< pagebreak >}}
+The biggest takeaways from our Exploratory Data Analysis is that most variables are fairly normally distributed due to the size of the data, and most variable are not correlated to each other, with one exception ('Age' and 'Marital')
+{{< pagebreak >}}
 
@@
 I chose to utilize the following models: the Multivariate Adaptive Regression Splines model (MARS), K-Nearest Neighbors model (KNN), Logistic Regression model, and RandomForest models to determine the best prediction model for this project.
+{{< pagebreak >}}
 
@@
 ```{r}
-#| label: fig-result999
+#| label: fig-result7
 #| fig-cap: "MARS Model example from MiniTab"
@@
 ```
-{{< pagebreak >}}
 
@@
 {{< pagebreak >}}
-### Background Regarding Performance Metrics
+### Background On Performance Metrics
 
@@
 
-**Recall:**\
+**Recall:(Sensitivity) **\
 The proportion of Positive cases that were correctly identified. This helps us understand how well we can classify positive cases specifically.\
+
+**Specificity:**\
+The proportion of Negative cases that were correctly identified.
 
@@
 The harmonic mean of Precision and Recall. This metric helps iron out extreme values\
+
+**ROC:**\
+The Receiver Operator Characteristic is a plotting technique that shows showing the threshold difference between Sensitivity (Recall) and Specificity.
+A model with good performance will have curve closer to the top left of the graph, whereas, a low-performing model will have a curve closer to the middle diagonal line.
+
+```{r}
+#| label: fig-result8
+#| fig-cap: "ROC diagram from DisplayR"
+#| echo: FALSE
+knitr::include_graphics(here("results","images","ROC_Example.jpg"))
+```
+
 {{< pagebreak >}}
@@
 #| label: tbl-resulttable1
+#| tbl-cap: "Performance Metrics for Numeric Unfactored Model Run"
 #| echo: FALSE
@@
 knitr::kable(resulttable1)
+#: My table {#tbl-resulttable1}
 ```
@@
 #| label: tbl-resulttable2
+#| tbl-cap: "Performance Metrics for Factored Model Run"
 #| echo: FALSE
@@
 In this run, the MARS model was the model to not identify a single positive class.
+
+@fig-result9 displays the ROC Curve for the first model run. Both ROC charts were identical, so only one is included here.
+
+```{r}
+#| label: fig-result9
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","ROC1.png"))
+```
+The ROC graph tells an even clearer story.
+While the RF model had a near perfect threshold trade-off between Sensitivity and Specificity, the other three models had very low ROC scores, again emphasizing poor predictive performance.
+
 {{< pagebreak >}}
@@
 
-@fig-result7 shows the Predictors that had the most impact on the Mars Model -- Numeric
+@fig-result10 shows the Predictors that had the most impact on the Mars Model -- Numeric
 
 ```{r}
-#| label: fig-result7
+#| label: fig-result10
 #| echo: FALSE
@@
 
-@fig-result9 shows the Predictors that had the most impact on the KNN Model--Numeric
+@fig-result11 shows the Predictors that had the most impact on the KNN Model--Numeric
 
 ```{r}
-#| label: fig-result9
+#| label: fig-result11
 #| echo: FALSE
@@
 
+## DownSampling Machine Learning Models
+In situations involving imbalanced response variables, there are different approaches available that can improve model performance.
+Kuhn and Johnson describe several different options including "Up-sampling" and "Down-sampling" (@kuhn2018).
+Up-sampling involves generating additional synthetic data in order to bring the totals of the response classes up to an equal amount.
+Down-sampling involves removing random observations until the the majority class is equal to the minority class (or classes).
+Since this project was already utilizing a fairly large dataset, I decided to perform Down-sampling in order to assist with the response variable imbalance that was being observed.
+The challenge with this approach is that there is less data to train with, but could lead to a greater overall model performance.
+
+I performed Down-Sampling on the training data set but kept the test data set intact as it was.
+It is recommended to allow the test data set to continue representing the natural data distribution in order for the most realistic prediction analysis
+
+### DownSampling Model Performance Metrics (Non-factored)
+
+```{r}
+#| label: tbl-resulttable3
+#| echo: FALSE
+resulttable3 = readRDS(here("results","tables","finaltable3.rds"))
+knitr::kable(resulttable3)
+```
+At first glance, it appears that the performance power on these models is less than the first models, but this is not the case.
+While Accuracy was lower among all models, Recall and F1 scores saw tremendous improvements.
+The Model that improved the most was clearly Logistic Regression which didn't show any Precision or F1 power initially, but now shows better Precision and F1 than the RandomForest model, which still showed to the be the superior model after this run.
+
+{{< pagebreak >}}
+
+### DownSampling Model Performance Metrics (Factored)
+
+```{r}
+#| label: tbl-resulttable4
+#| echo: FALSE
+resulttable4 = readRDS(here("results","tables","finaltable4.rds"))
+knitr::kable(resulttable4)
+```
+We can observe similar changes in the second Down-sampled run.
+In the Factored-Down-Sampled model runs, we also see that it is harder to identify a single model of best performance.
+The MARS model showed the best, Accuracy and Specificity, while the RandomForest model showed the best Sensitivity and Precision.
+The Logistic Regression model was not far behind in these categories and managed to approximately tie three ways for the F1 metric, along with the MARS Model and RandomForest Model.
+{{< pagebreak >}}
+
+@tbl-resulttable15 displays the ROC Curve for the Down-Sampled model runs. Again, both ROC charts were identical, so only one is included here.
+
+```{r}
+#| label: tbl-resulttable15
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","ROC3.png"))
+```
+We can see much better results on the ROC model compared to the first runs.
+The RandomForest again performs nearly perfectly, but all three remaining models show some promise in terms of threshold trade-off, with the MARS model leading these three.
+
+{{< pagebreak >}}
+
+### Variables of Importance for Down-Sampled Models
+
+@fig-result16 shows the Predictors that had the most impact on the Down-Sampled Mars Model -- Factored
+
+```{r}
+#| label: fig-result16
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","DSimportantM2.png"))
+```
+Both MARS models highlighted essentially the same variables of importance, with added detail coming from the factored model.
+As mentioned previously, Housing2 indicates that the person does have a housing loan and Loan2 indicates that the person has personal loan.
+Education3 indicates that the person has attained a tertiary level of education.
+Marital 2 indicates the person is married.
+{{< pagebreak >}}
+
+@fig-result17 shows the Predictors that had the most impact on the Down-Sampled KNN Models (both models highlighted the same variables)
+
+```{r}
+#| label: fig-result17
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","DSimportantKNN1.png"))
+```
+The variables highlighted here are the same that were highlighted in the original KNN models with the only exception being the Marital variable moved up one spot on this list.
+{{< pagebreak >}}
+
+@fig-result18 shows the Predictors that had the most impact on the Down-Sampled Logistic Regression Model -- Factored
+
+```{r}
+#| label: fig-result18
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","DSimportantLog2.png"))
+```
+These are practically the same variables highlighted in the original Logistic Regression Model, except while the 'Education3' variable was the third most important variable on the original model, it is nowhere to be found on this new list.
+In contrast, the 'Job3' (Technician) variable makes an appearance as having a substantial impact on prediction power.
+{{< pagebreak >}}
+
+@fig-result19 shows the Predictors that had the most impact on the Down-Sampled Logistic Regression Model -- Factored
+
+```{r}
+#| label: fig-result19
+#| echo: FALSE
+knitr::include_graphics(here("results","figures","DSimportantRF1.png"))
+```
+While the two Down-Sampled models were fairly close in highlighting variables, there was quite a bit of difference between these results and the original RandomForest model.
+Where the original highlighted, Age, Housing, Day and Balance, the new models highlighted Housing, Age, Balance, and Loan.
+{{< pagebreak >}}
+
 # Discussion
@@
 
-  Despite these findings, the models were able to produce which variables provided the most weight towards the prediction power.
+  Using a Down-Sampling approach to alleviate the imbalance issue resulted in models with greater predictive power and gave more clarity on which variables (and in some cases, sub-variables) contributed the most towards correctly predicting the response variable.
+
+  All four models provided insights for this objective.
 The Housing variable was listed as the most important variable by every model with the exception of the Mars-Numeric model.
@@
 
-{{< pagebreak >}}
-
 # References

### 2024-08-05 code/analysis-code/statistical-analysis - my version.R

@@ -1,390 +1,0 @@
-###############################
-
-#required libraries
-library(ggplot2) #for plotting
-library(broom) #for cleaning up output from lm()
-library(here) #for data loading/saving
-library(earth)
-library(tidyr)
-library(dplyr)
-library(caret)
-library(knitr)
-library(randomForest)
-library(Metrics)
-
-
-#loading data. 
-data_location <- here::here("data","processed-data","processeddata4.rds")
-mydata <- readRDS(data_location)
-summary(mydata)
-
-#Basic Statistical Analysis
-#Examining a logistic regression test with all variables of interest
-
-logitFull = glm(y_termSubscribed~., data = mydata, family = "binomial")
-summary(logitFull)
-
-lrtable <- broom::tidy(logitFull)
-fullmodel = here("results", "tables", "fullmodel.rds")
-saveRDS(lrtable, file = fullmodel)
-
-#All Predictors are Significant
-
-#Full Data Analysis
-
-
-######################################
-#Machine Learning Models Part 1
-######################################
-#Using Numeric Unfactored Variables
-############################
-
-
-#Preprocessing 2 True Numeric Variables
-set.seed(24)
-mydataPP = preProcess(mydata[7:8], method = c("scale", "center", "BoxCox"))
-mydata[7:8] = predict(mydataPP, mydata[7:8])
-
-
-#Splitting the data. 80% for training, 20% test.
-
-set.seed(21)
-split1 = sample(c(rep(0, 0.8 * nrow(mydata)), rep(1, 0.2*nrow(mydata))))
-
-
-train_data = mydata[split1 == 0, ]
-test_data = mydata[split1 == 1, ]
-
-train_y = train_data$y_termSubscribed
-test_y = test_data$y_termSubscribed
-
-
-train_x = train_data %>% 
-  select(!y_termSubscribed)
-
-test_x = test_data %>% 
-  select(!y_termSubscribed)
-
-
-
-#Setting Tran Control
-ctrl <- trainControl(method = "cv")
-
-#################
-#Modeling Section
-
-# MARS model
-set.seed(21)
-marsmodel = train(y_termSubscribed~., data=train_data, 
-                  method = "earth",
[…]
